---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Load required libraries
```{r loadlibraries}
library(ggplot2)
library(dplyr)
library(simputation)
```

## Loading and preprocessing the data

```{r readfile}
movements<-read.csv("activity.csv",header = TRUE,sep = ",")
# Convert char to Date
movements$date<-as.Date(movements$date,format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

    1. Calculate the total number of steps taken per day

    2. Make a histogram of the total number of steps taken each day

    3. Calculate and report the mean and median of the total number of steps taken per day

```{r totalsteps, results=FALSE}
totalStepsPerDay<-aggregate(movements$steps, by=list(movements$date), FUN=sum, na.rm=TRUE)
colnames(totalStepsPerDay) <- c('Date', 'TotalSteps')
totalStepsMean<-mean(totalStepsPerDay$TotalSteps)
totalStepsMedian<-median(totalStepsPerDay$TotalSteps)
hist(totalStepsPerDay$TotalSteps,main="Total Daily Steps",xlab="")
```

The median number of steps taken per day is `r totalStepsMedian`  

The mean number of steps taken per day is `r totalStepsMean`  


## What is the average daily activity pattern?
```{r averagesteps, result=FALSE}
dailyActivity <- movements %>%
  na.omit() %>%
  group_by(date, interval) %>%
  dplyr::summarize(AvgSteps = mean(steps)) %>% 
  as.data.frame()

pl<-ggplot(data=dailyActivity,aes(date,AvgSteps))
pl+geom_line()
q1<-dailyActivity[which.max(dailyActivity$AvgSteps),]
```

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is interval `r q1[2]` on `r q1[1]` with a value of `r q1[3]` steps  


## Imputing missing values

```{r missingvalues, result=FALSE}
missingSteps<-sum(is.na(movements$steps))
missingDate<-sum(is.na(movements$date))
missinginterval<-sum(is.na(movements$interval))
```

There are `r missingSteps` missing values for Steps,  `r missingDate` missing values for Date and `r missinginterval` missing values for interval  

### Create a new dataset that is equal to the original dataset but with the missing data filled in


```{r imputemissingvalues, result=TRUE}

cleanMovements<-impute_lm(movements,steps ~ date)
cleanStepsPerDay<-aggregate(cleanMovements$steps, by=list(cleanMovements$date),FUN=sum)
colnames(cleanStepsPerDay) <- c('Date', 'DailySteps')
hist(cleanStepsPerDay$DailySteps,main="Total Daily Steps",xlab="")
cleanTotalStepsMean<-mean(cleanStepsPerDay$DailySteps)
cleanTotalStepsMedian<-median(cleanStepsPerDay$DailySteps)
diffMean<-abs(totalStepsMean - cleanTotalStepsMean)
diffMedian<-abs(totalStepsMedian - cleanTotalStepsMedian)
```
The median number of steps taken per day is `r cleanTotalStepsMedian`  

The mean number of steps taken per day is `r cleanTotalStepsMean`  

The difference in the mean between the two data set is `r diffMean`  

The difference in the median between the two data set is `r diffMedian`  

The median frequency of the steps has increaded from 25 to 35


## Are there differences in activity patterns between weekdays and weekends?
