---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Load required libraries
```{r loadlibraries}
library(ggplot2)
library(dplyr)
```

## Loading and preprocessing the data

```{r readfile}
movements<-read.csv("activity.csv",header = TRUE,sep = ",")
movements$Dates<-as.Date(movements$date,format="%Y-%m-%d")
```


## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

    1. Calculate the total number of steps taken per day

    2. Make a histogram of the total number of steps taken each day

    3. Calculate and report the mean and median of the total number of steps taken per day

```{r totalsteps, results=FALSE}
totalStepsPerDay<-aggregate(movements$steps, by=list(movements$Dates), FUN=sum, na.rm=TRUE)
colnames(totalStepsPerDay) <- c('Date', 'TotalSteps')
totalStepsMean<-mean(totalStepsPerDay$TotalSteps)
totalStepsMedian<-median(totalStepsPerDay$TotalSteps)
hist(totalStepsPerDay$TotalSteps,main="Total Daily Steps",xlab="")
```

#### The media number of steps taken per day is `r totalStepsMedian`  

#### The mean number of steps taken per day is `r totalStepsMean`  


## What is the average daily activity pattern?
```{r averagesteps, result=FALSE}
dailyActivity <- movements %>%
  na.omit() %>%
  group_by(Dates, interval) %>%
  dplyr::summarize(AvgSteps = mean(steps)) %>% 
  as.data.frame()

pl<-ggplot(data=dailyActivity,aes(Dates,AvgSteps))
pl+geom_line()
q1<-dailyActivity[which.max(dailyActivity$AvgSteps),]
```

### The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is interval `r q1[2]` on `r q1[1]` with a value of `r q1[3]` steps  


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
